% !TeX root = ../thuthesis-example.tex

\begin{resume}

  \section*{Education}

  2022 --- 2024: \textbf{Tsinghua University}, Computer Science and Technology department, Master’s Program ``Advanced Computing Program''.
  Supervisor --- Associate Professor X. Hu, thesis on physical adversarial attacks.

  2018 --- 2022: \textbf{National Research University Higher School of Economics}, Faculty of Computer Science, Bachelor’s Program ``Applied Mathematics and Information Science''.
  Supervisor --- Associate Professor M. Rakhuba, thesis on theoretical ML. 
  Minor in Bioinformatics.

  \section*{Work Experience}
  2022 --- 2024: \textbf{Tsinghua} Research Intern in the Tsinghua Laboratory of Brain and Intelligence.

  2021 --- 2024: \textbf{HSE, Moscow}: Teaching Assistant to Researcher E. Lobacheva.
  
  2021 --- 2022: \textbf{HSE, Moscow} Research Intern in the Laboratory of Stochastic Algorithms and High-Dimensional Inference.
  
  2021: \textbf{HSE, Moscow} Research Intern in the Laboratory of Theoretical Computer Science.

  \section*{Achievements}

  \subsection*{Achievements}

  \begin{achievements}
    \item \href{https://openreview.net/forum?id=T5TtjbhlAZH}{``Towards Practical Control of Singular Values of Convolutional Layers''}, \textbf{NeurIPS 2022}.
    \item First place in the ``Student Research Paper Competition'' for bachelor's students held by \textbf{HSE University}, 2021.
  \end{achievements}

  \subsection*{Scholarships}

  \begin{achievements}
    \item 2022 --- 2024: Tsinghua Scholarship for Graduate Students.
    \item 2022: Academic Scholarship \& Travel Grant for winning the ``Student Research Paper Competition''.
    \item 2018 --- 2022: Russian State Academic Scholarship.
    \item 2018 --- 2020: Moscow Government Scholarship for Distinguished Achievements in Education.
  \end{achievements}

  \section*{Additional Information}

  \paragraph{Languages} Russian (native), English (C1), Chinese (Intermediate).

  \paragraph{Technical Skills} Python, C/C++, \LaTeX, Java, bash.

  \paragraph{Other Skills} Arts, Translation, Linguistics.

\end{resume}
